---
title: "Interactive Timelines of Nutrition, Food and Health"
author: "Cathie Wotecki, Brandon Kramer, Vicki Lancaster and Sam Cohen"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls())
for (pkg in c("ggplot2", "scales", "lubridate", "googlesheets", "timevis", "googleVis", "tidyverse", "webshot")) {library(pkg, character.only = TRUE)}
```

Next, we decided to establish a basic timeline of major events in the nutrition sciences over time. We decided to break this down into three distinct timelines based on science, politics and the social.

**Here is the science timeline...**

```{r science timeline, warning=FALSE, message=FALSE, fig.width=9.5, fig.height=2}
setwd("~/Documents/Nutrition/impacts-and-echoes/data")
timeline_data <- read_csv("interactive_timeline_data - final_timeline.csv")

science_timeline <- timeline_data %>% 
  filter(domain == "Science" )

timevis(science_timeline)
```


**Here is the political timeline...** 

```{r policy timeline, warning=FALSE, message=FALSE, fig.width=9.5, fig.height=3}
policy_timeline <- timeline_data %>% 
   filter(domain == "Policy")

timevis(policy_timeline)
```

**Here is the social timeline...** 

```{r social events timeline, fig.height=3, fig.width=9.5, message=FALSE, warning=FALSE}
social_timeline <- timeline_data %>% 
  filter(domain == "Social")

timevis(social_timeline)
```

**And here is a summarized or selective timeline of events...** 

```{r general timeline, fig.width=9.5, fig.height=10.5}
summarized_timeline <- timeline_data %>% 
  filter(final == "final")  
plot <- timevis(summarized_timeline, 
        groups = data.frame(id = 1:6, content = c("Social", "Food", "Science", 
                                                  "Health", "Policy", "Conference")))
plot 
```

```{r, fig.width=7, fig.height=9, include = FALSE}
setwd("~/Documents/Nutrition/impacts-and-echoes")
htmlwidgets::saveWidget(plot, "timeline.html", selfcontained = F)
```

```{r, fig.width=7, fig.height=9, include = FALSE}
setwd("~/Documents/Nutrition/impacts-and-echoes/docs")
webshot("https://brandonleekramer.github.io/impacts-and-echoes/timeline.html", file="timeline.pdf")
```

We also worked on converting this interactive timeline to an .svg but this was not successful. 

















